Appendix 1: sdmTMB statistical model description

### 2021-12-21 doc/appendix-model.Rmd
Edit

* Greek symbols for parameters

* Bold symbols for vectors or matrices

* $\phi$ for all distribution dispersion parameters for consistency with the code

* $\mathbb{E}[y]$ to define the expected value (mean) of variable $y$

* $\mathrm{Var}[y]$ to define the expected variance of the variable $y$

* An $^*$ superscript to represent interpolated or projected values as opposed to values at knot locations (e.g., $\omega$ vs. $\omega^*$).

```{r, results='asis'}
symbols <- tibble::tribble(
  ~Symbol, ~Description,
  "$y$", "Observed response data",
  "$\\mu$", "Mean",
  "$\\phi$", "A dispersion parameter for a distribution",
  "$f$", "Link function",
  "$f^{-1}$", "Inverse link function",
  "$\\boldsymbol{\\beta}$", "Parameter vector",
  "$\\boldsymbol{X}$", "A predictor model matrix",
  "$O_{\\s, t}$", "An offset variable at point $\\s$ and time $t$",

@@

```{r}
subscripts <- tibble::tribble(
  ~Symbol, ~Description,
  "$\\s$", "Index for space; a vector of x and y coordinates",
  "$t$", "Index for time",
  "$g$", "Group"
)
knitr::kable(subscripts, caption = "Subscript notation", escape = FALSE, format = "latex", booktabs = TRUE)
```

@@

@@

where $\alpha_g$ is an example optional "random" intercept---an intercept
centered on zero that varies by level $g$ and is constrained by
$\sigma_\alpha$. This is defined by the `formula` argument via the `(1 | g)`
syntax as in lme4 or glmmTMB. There can be multiple random intercepts. 
E.g., `(1 | g) + (1 | h)`, in which case they are assumed independent and uncorrelated from each other.

@@

[Source code](https://kaskr.github.io/adcomp/tweedie_8cpp_source.html) as in [cplm](https://CRAN.R-project.org/package=cplm) package. <https://link.springer.com/article/10.1007/s11222-012-9343-7>

>>>>

* Greek symbols for parameters,

* the Latin/Roman alphabet for data,

* bold symbols for vectors or matrices,

* $\phi$ for all distribution dispersion parameters for consistency with the code,

* $\mathbb{E}[y]$ to define the expected value (mean) of variable $y$,

* $\mathrm{Var}[y]$ to define the expected variance of the variable $y$, and

* a $^*$ superscript to represent interpolated or projected values as opposed to values at knot locations (e.g., $\omega$ vs. $\omega^*$).

We include tables of all major indices (Table \@ref(tab:indices)) and symbols (Table \@ref(tab:symbols)).

```{r indices}
library(magrittr)
subscripts <- tibble::tribble(
  ~Symbol, ~Description,
  "$\\s$", "Index for space; a vector of x and y coordinates",
  "$t$", "Index for time",
  "$g$", "Group"
)
knitr::kable(subscripts, caption = "Subscript notation", escape = FALSE, format = "latex", booktabs = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "hold_position")
```

```{r symbols, results='asis'}
symbols <- tibble::tribble(
  ~Symbol, ~Description,
  "$y$", "Observed response data",
  "$\\mu$", "Mean",
  "$\\phi$", "A dispersion parameter for a distribution",
  "$f$", "Link function",
  "$f^{-1}$", "Inverse link function",
  "$\\boldsymbol{\\beta}$", "Parameter vector",
  "$\\boldsymbol{X}$", "A predictor model matrix",
  "$O_{\\s, t}$", "An offset variable at point $\\s$ and time $t$",

@@

\clearpage

@@

The random fields are parameterized internally with a sparse precision matrix ($\boldsymbol{Q}_\zeta$):

$$
\boldsymbol{\zeta} \sim \mathrm{MVNormal}\left(\boldsymbol{0}, \sigma_\zeta^2 \boldsymbol{Q}^{-1}_\zeta\right).
$$

@@

where $\alpha_g$ is an example optional "random" intercept---an intercept
with mean zero that varies by level $g$ and is constrained by
$\sigma_\alpha$. This is defined by the `formula` argument via the `(1 | g)`
syntax as in lme4 or glmmTMB. There can be multiple random intercepts. 
E.g., `(1 | g) + (1 | h)`, in which case they are assumed independent and uncorrelated from each other.

@@

[Source code](https://kaskr.github.io/adcomp/tweedie_8cpp_source.html) as in the [cplm](https://CRAN.R-project.org/package=cplm) package [@zhang2013a].
